---
title: "llm_segmentation_analysis"
output: html_document
date: "2024-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggdist)
library(DescTools)
library(emmeans)
library(lmerTest)
library(scales)
library(viridis)
library(ggsci)
library(broom)
library(ggbeeswarm)
library(effectsize)
library(readxl)
library(patchwork)
library(report)
source('agreement.R')
```

### Import Data

```{r}
data_path = 'PATH_TO/'
llm_data = read_csv(paste0(data_path, 'llm_segmentation_data.csv')) %>% 
  mutate(story_id = as.factor(story_id),
  			 temperature = as.factor(temperature))

llm_series = read_csv(paste0(data_path, 'llm_time_word_series.csv')) %>% 
	  mutate(story_id = as.factor(story_id),
  			 temperature = as.factor(temperature))

visual_data = read_csv(paste0(data_path,'visual_segmentation_data.csv')) %>% 
  mutate(story_id = as.factor(story_id))

visual_series = read_csv(paste0(data_path, 'visual_time_word_data.csv')) %>% 
  mutate(story_id = as.factor(story_id))
```

### Constants

```{r}
# Total Words
story1_word = 1669
story2_word = 1521
story3_word = 1740

# Colours for Plots 
colours = c('#3895B6', '#41D3CB', '#38B686', '#5456BC')
```

### Format Functions

```{r}
format_value <- function(data, column, digits = 3, use_round = FALSE) {
  if (use_round) {
    data %>% mutate(column = round(column, digits))
  } else {
    data %>% mutate(column = format(column, digits = digits))
  }
}

format_value = function(x) mutate(x)
```

## Number of Events

```{r}
num_events = llm_data %>% 
  rename(group = temperature,
         word_number = word_number) %>% 
  rbind(., visual_data) %>% 
  group_by(subject, story_id, group) %>% 
  summarize(num_events = n()) %>% 
  mutate(scaled_num_events = num_events / case_when(story_id == 1 ~ story1_word,
                                                    story_id == 2 ~ story2_word,
                                                    story_id == 3 ~ story3_word) * 1000)
```

### Statistics

```{r}
model = num_events %>% 
	lmer(num_events ~ group + (1 | story_id), data=.)

anova(model)
emmeans(model, ~group) %>% 
  pairs()
eta_squared(model)

```

### Plot

```{r}
p1 = ggplot(num_events, aes(x=group, y=scaled_num_events, fill=group)) +
	stat_summary(fun.data = mean_sdl, geom = 'bar') +
	geom_quasirandom(alpha = 0.1, width = 0.2) +
	stat_summary(fun.data = mean_se, geom = 'errorbar', width = 0.2, linewidth = 1) +
	scale_x_discrete(labels = c('0', '0.5', '1', 'Human')) +
	scale_fill_manual(values = colours) +
  theme_ggdist() +
  labs(x='Condition', y='Num Events', title = 'Number of Events') +
  theme(axis.title = element_text(face='bold'), legend.position = 'none')
p1
```

## LLaMA Agreement Index

```{r}
grouped_data = llm_series %>% 
  group_by(story_id, temperature) %>% 
  group_split()

agreement_llm = tibble()

for (df in grouped_data){
  subjects = unique(df$subject)
  story = unique(df$story_id)
  temperature = unique(df$temperature)
  
  print(story)
  print(temperature)
  
  # Assign Story Duration
  story_dur = case_when(
    story == 1 ~ story1_word,
    story == 2 ~ story2_word,
    story == 3 ~ story3_word
  )
  
  ag_index = c()
  for (subj in subjects){
    
    print(subj)
    # Subset data
    dat = df %>% 
      filter(subject == subj)
    
    # Calculate Agreement Index
    sample = dat$responses
    group = tapply(df$responses[df$subject != subj], 
                   df$word_series[df$subject != subj], mean)
      
    
    agreement = agreement_index(sample, group)
    
    ag_index = rbind(ag_index, agreement)
  }
  
  temp = tibble(subject = unique(df$subject), 
                story_id = story, 
                temperature = temperature, 
                agreement_index = ag_index[,1])
  
  agreement_llm = rbind(agreement_llm, temp)
}
```

## Human Agreement Index

```{r}
grouped_data = visual_series %>% 
  group_by(story_id) %>% 
  group_split()

agreement_visual = tibble()

for (df in grouped_data){
  subjects = unique(df$subject)
  story = unique(df$story_id)
  
  print(story)
  
  # Assign Story Duration
  story_dur = case_when(
    story == 1 ~ story1_word,
    story == 2 ~ story2_word,
    story == 3 ~ story3_word
  )
  
  ag_index = c()
  for (subj in subjects){
    
    print(subj)
    # Subset data
    dat = df %>% 
      filter(subject == subj)
    
    # Calculate Agreement Index
    sample = dat$responses
    group = tapply(df$responses[df$subject != subj], 
                   df$word_number[df$subject != subj], mean)
      
    
    agreement = agreement_index(sample, group)
    
    ag_index = rbind(ag_index, agreement)
  }
  
  temp = tibble(subject = unique(df$subject), 
                story_id = story, 
                agreement_index = ag_index[,1])
  
  agreement_visual = rbind(agreement_visual, temp)
}
```

```{r}
# Compile Data
agreement = agreement_llm %>% 
	rename(condition = temperature) %>% 
	rbind(agreement_visual %>% 
					mutate(condition = 'human'))
```

### Statistics

```{r}
agreement_model = agreement %>% 
  lmer(agreement_index ~ condition + (1 | subject) + (1 | story_id), data = .) 

agreement_model %>% 
  anova()

agreement_model %>% 
  eta_squared()
  
agreement_model %>% 
  emmeans(~ condition) %>% 
  pairs()
```

### Plot

```{r}
p2 = ggplot(agreement, aes(condition, agreement_index, fill = condition)) +
  stat_summary(fun.data = mean_se, geom = 'bar') +
	scale_x_discrete(labels = c('0', '0.5', '1', 'Human')) +
	scale_fill_manual(values = colours) +
  geom_quasirandom(alpha = 0.1, varwidth = TRUE) +
	stat_summary(fun.data = mean_se, geom = 'errorbar', width = 0.2, linewidth = 1) +
	labs(x = 'Condition', y = 'Agreement Index', title='Agreement Index Within Group') +
	theme_ggdist() +
	theme(axis.title = element_text(face = 'bold'),
				legend.position = 'none')
p2
```

## Human-LLaMA Agreement Index

Human subjects compared to the LLaMA produced group event series.

```{r}
grouped_data = llm_series %>% 
  filter(word_series != 0) %>% 
  group_by(story_id, temperature) %>% 
  group_split()

agreement_llm_human = tibble()

visual_subjects = visual_data %>% 
  select(subject) %>% 
  distinct() %>% 
  pull()

for (df in grouped_data){
  subjects = unique(df$subject)
  story = unique(df$story_id)
  temperature = unique(df$temperature)
  
  # Assign Story Duration
  story_dur = case_when(
    story == 1 ~ story1_word,
    story == 2 ~ story2_word,
    story == 3 ~ story3_word
  )
  
  ag_index = c()
  group = tapply(df$responses, df$word_series, mean)
  
  agreement_llm_human = visual_series %>% 
    filter(story_id == story) %>% 
    group_by(subject, story_id) %>% 
    summarise(temperature = temperature,
              agreement_index = agreement_index(responses, group), .groups = 'drop') %>% 
    rbind(agreement_llm_human, .)
}
```

### Statistics

```{r}
agreement_model = agreement_llm_human %>% 
  mutate(across(c(subject, story_id, temperature), as.factor)) %>% 
  lmer(agreement_index ~ temperature + (1 | subject) + (1 | story_id), data=.)

agreement_model %>% 
  anova()

agreement_model %>%
  eta_squared()
  
agreement_model %>% 
  emmeans(~ temperature) %>% 
  pairs()
```

### Plot

```{r}
p3 = agreement_llm_human %>% 
	mutate(temperature = as.factor(temperature)) %>% 
  ggplot(aes(temperature, agreement_index, fill = temperature)) +
	stat_summary(fun.data = mean_sdl, geom = 'bar') +
  geom_line(aes(group=interaction(subject,story_id)), alpha = 0.1) +
	stat_summary(fun.data = mean_se, geom = 'errorbar', width = 0.2, linewidth = 1) +
	labs(x = 'Temperature', y = 'Agreement Index') +
	scale_fill_manual(values = colours) +
	theme_ggdist() +
	theme(axis.title = element_text(face = 'bold'),
				legend.position = 'none')
p3
```

## Histogram

```{r}
compiled_data = llm_series %>% 
	filter(word_series != 0) %>% 
	rename(word_number = word_series) %>% 
	rbind(visual_series %>% 
	        select(!c(word, smoothed_responses)) %>% 
	        mutate(temperature = 'human')) %>% 
	rename(condition = temperature)
```

### Plot

```{r}
p4 = compiled_data %>% 
	filter(responses == 1, story_id == 1) %>% 
	ggplot(aes(x=word_number, colour = condition, fill = condition)) +
	geom_histogram(aes(y = after_stat(count)/20), binwidth = 1) +
	facet_wrap(~ condition, ncol=1) +
	scale_y_continuous(breaks = c(0, 0.5, 1)) +
	scale_colour_manual(values = colours, 
	                    labels = c('0'= '0', '0.5' = '0.5', '1' = '1', 'human' = 'Human')) + 
	scale_fill_manual(values = colours, 
	                  labels = c('0'= '0', '0.5' = '0.5', '1' = '1', 'human' = 'Human')) +
	labs(x = 'Word Number', y = 'Proportion of Participants', 
	     title='Event Boundaries Identified') +
	theme_ggdist() +
  theme(legend.position = 'right', 
        legend.title = element_blank(),
        strip.text = element_blank(),
        axis.title = element_text(face='bold'))
p4
```

## Amplitude Average at Match

```{r}
common_index = function(dataframe, group1, group2){
  common_index = dataframe %>% 
    filter(condition == group1) %>% 
    semi_join(dataframe %>% 
                filter(condition == group2),
              by = 'word_series') %>% 
    pull(word_series)
  
  uncommon_index = dataframe %>% 
    filter(condition == group1) %>% 
    anti_join(dataframe %>% 
                filter(condition == group2),
              by = 'word_series') %>% 
    pull(word_series)
  
  list = list(common_index = common_index, uncommon_index = uncommon_index)
  
  return(list)
}

lighten_colour = function(colour, factor=0.6) {
  return(colorspace::lighten(colour, amount=factor))
}

all_colours = c(colours, sapply(colours, lighten_colour, factor=0.5))
names(all_colours) <- c('0_Match', '0.5_Match', '1_Match', '2_Match',
                       '0_No Match', '0.5_No Match', '1_No Match', '2_No Match')

names(all_colours) = c('0_common', '0.5_common', '1_common', '2_common',
                      '0_uncommon', '0.5_uncommon', '1_uncommon', '2_uncommon')
```

```{r}
# Compile and Average Word Series Data
avg_human_series = visual_series %>% 
  group_by(story_id, word_number) %>%
  summarize(avg_responses = mean(responses)) %>% 
  rename(word_series = word_number) %>% 
  mutate(condition = 'human')

avg_llm_series = llm_series %>% 
  group_by(story_id, word_series, temperature) %>% 
  summarize(avg_responses = mean(responses)) %>% 
  filter(word_series != 0) %>% 
  arrange(temperature) %>% 
  rename(condition = temperature)

avg_series = avg_human_series %>% 
  rbind(avg_llm_series) %>% 
  filter(avg_responses != 0)


avg_amp_human = tibble()
amp_human = tibble()
for (story in c(1,2,3)){
  for (temp in c(0,0.5,1)){
    index = avg_series %>% 
      filter(story_id == story) %>% 
      common_index(dataframe=., group1 = 'human', group2 = as.character(temp))
    
    average_amplitude_common = avg_series %>% 
      filter(condition == 'human', story_id == story) %>%
      filter(word_series %in% index$common_index) %>% 
      summarize(average_amplitude = mean(avg_responses)) %>% 
      pull()
    
    amplitude_common = avg_series %>% 
      filter(condition == 'human', story_id == story) %>%
      filter(word_series %in% index$common_index) %>% 
      mutate(temperature = temp,
             amp_type = 'common')
      
    average_amplitude_uncommon = avg_series %>% 
      filter(condition == 'human', story_id == story) %>%
      filter(word_series %in% index$uncommon_index) %>% 
      summarize(average_amplitude = mean(avg_responses)) %>% 
      pull()
    
    amplitude_uncommon = avg_series %>% 
      filter(condition == 'human', story_id == story) %>%
      filter(word_series %in% index$uncommon_index) %>% 
      mutate(temperature = temp,
             amp_type = 'uncommon')
    
    avg_amp_human = tibble(story_id = story, temperature = temp, 
                average_amplitude_common = average_amplitude_common,
                average_amplitude_uncommon = average_amplitude_uncommon) %>% 
      rbind(avg_amp_human)
    
    amp_human = rbind(amplitude_common, amplitude_uncommon) %>% 
      rbind(amp_human)
    
  }
}
```

### Statistics

```{r}
amp_model = amp_human %>% 
  mutate(across(c(temperature, condition, amp_type), as.factor)) %>%
  lmer(avg_responses ~ temperature * amp_type + (1 | story_id), data = .) 

amp_model %>%   
  anova() 

amp_model %>%   
  eta_squared()

amp_model %>% 
  emmeans(~ temperature * amp_type) %>% 
  pairs() 
```

### Plot

```{r}
p5 = amp_human %>% 
  mutate(temp_type = paste(temperature, amp_type, sep = '_')) %>% 
  ggplot(aes(x=as.factor(temperature), y=avg_responses, fill=temp_type)) +
  stat_summary(fun.data = mean_sdl, geom = 'bar', position = 'dodge') +
  stat_summary(fun.data = mean_se, geom = 'errorbar', 
               position = position_dodge(width = 0.9), linewidth = 1, width = 0.2) +
  scale_fill_manual(values = all_colours, labels = c('Match', 'No Match')) +
  labs(x = 'Temperature', y = 'Proportion of Participants', title='Event Match Proportion') +
  theme_ggdist() +
  theme(axis.title = element_text(face = 'bold'), 
        legend.position = 'none')
p5
```

## Number of Peaks Match

Proportion of peaks matching a split-half sample of human raters to a second split-half sample of given condition.

```{r}
proportion_match_split = function(dataframe, group1, group2, iter){
  subjects = unique(dataframe$subject)

  proportion_match = tibble()
  for (i in 1:iter){
    subj1 = sample(subjects, length(subjects) / 2)
    subj2 = setdiff(subjects, subj1)
    
    series1 = dataframe %>% 
      filter(condition == group1, subject %in% subj1) %>% 
      group_by(word_series, condition) %>% 
      summarize(avg_responses = mean(responses), .groups = 'drop_last')
    
    num_peaks1 = series1 %>% 
      filter(avg_responses != 0) %>% 
      pull(word_series) %>% 
      length()
    
    series2 = dataframe %>% 
      filter(condition == group2, subject %in% subj2) %>% 
      group_by(word_series, condition) %>% 
      summarize(avg_responses = mean(responses), .groups = 'drop_last')
    
    num_peaks2 = series2 %>% 
      filter(avg_responses != 0) %>% 
      pull(word_series) %>% 
      length()
    
    if (group1 == group2){
      new_group1 = paste(group1, '1', sep = '')
      new_group2 = paste(group2, '2', sep = '')
      
      series1 = series1 %>% 
        mutate(condition = new_group1)
      series2 = series2 %>% 
        mutate(condition = new_group2)
      
      longer_group = ifelse(num_peaks1 >= num_peaks2, new_group1, new_group2)
      shorter_group = setdiff(c(new_group1, new_group2), longer_group)
    } else {
    longer_group = ifelse(num_peaks1 >= num_peaks2, group1, group2)
    shorter_group = setdiff(c(group1, group2), longer_group)
    }
    
    index = rbind(series1, series2) %>% 
      filter(avg_responses != 0) %>% 
      common_index(dataframe=., group1=shorter_group, group2=longer_group)
    
    proportion = length(index$common_index) / (length(index$common_index) + length(index$uncommon_index))
    
    proportion_match = tibble(condition1 = group1, condition2 = group2, iteration = i, 
                              proportion = proportion) %>% 
      rbind(proportion_match)
    
  }
  
  return(proportion_match)
}
```

```{r}
# Compile Word Series Data
compiled_series = visual_series %>% 
  select(subject, story_id, word_number, responses) %>% 
  rename(word_series = word_number) %>% 
  mutate(condition = 'human') %>% 
  rbind(llm_series %>% 
          filter(word_series != 0) %>% 
          rename(condition = temperature))

proportion_match = tibble()
for (i in unique(avg_series$condition)){
  for (story in unique(avg_series$story_id)){
    cat('Condition', i, 'Story', story)
    proportion_match = compiled_series %>% 
      filter(story_id == story) %>% 
      proportion_match_split(dataframe=., group1='human', group2=i, iter=100) %>% 
      mutate(story_id = story) %>% 
      rbind(proportion_match)
    
  }
}
```

### Statistics

```{r}
match_model = proportion_match %>% 
  mutate(across(c(condition2, story_id), as.factor)) %>% 
  lmer(proportion ~ condition2 + (1 | story_id), data=.)

match_model %>% 
  anova()

match_model %>% 
  eta_squared()

match_model %>% 
  emmeans(~ condition2) %>% 
  pairs()

```

### Plot

```{r}
p6 = ggplot(proportion_match, aes(x=condition2, y=proportion, fill=condition2)) +
  stat_summary(fun.data = 'mean_se', geom = 'bar') +
  # geom_jitter(alpha=0.2, width =0.2) +
  stat_summary(fun.data = 'mean_se', geom = 'errorbar', width=0.2, size = 1) +
  scale_x_discrete(labels = c('0'= '0', '0.5' = '0.5', '1' = '1', 'human' = 'Human')) +
  labs(x = 'Condition', y = 'Proportion Match', title='Split-Half Event Match') +
  scale_fill_manual(values = colours) +
  theme_ggdist() +
  theme(axis.title = element_text(face = 'bold'), legend.position = 'none')

p6
```
